### 2015-07-21 text/ms.Rmd
Incorporate more of Jessica's comments

<!--Jessicas Comments: 
1. Massive improvement in abstract and introduction. I really like the strucutre and how you explain the models. makes sense to people who aren't familiar with the problems around datalimited fisheries, and the ideas and need for ensemble models. This is an amazing paper - quite complicated, but well explained to show people why it matters and how they can use it. 
2. If I remember correctly, Cons Biol doesn't like having the text wrapped around between two lines with hyphens as the PDF shows. Not sure how to fix that, but just something to note. 
3. is it normal to describe the results and findings at the end of the introduction? i see it sometimes, but not sure if Cons BIol like it. If not, you could put that final paragraph of introduction at start of discussion.
4. First paragraph in Methods describing simulations line 113 in pdf - you may want to describe the recrutiment variability a bit more. I know its not an important detail, but because its in the first paragraph of methods, it might turn people off reading further. Most people wouldn't understand what 'multiplcative' (including me) is or may not know what 'process error/noise' refers to. 
5. First paragraph in Methods last sentence - needed a bit of rewording because as was, i got quite confused. hopefully this new sentence makes sense and doesnt change the meaning. 
6. Fix up Simulation FLR package
7. Update Version xx downloaded on xx RAM legacy database
8. line 125 in pdf - Ram legacy database - how many stocks didn't converge? readers might want to know how many stocks you threw away.
9. make datalimited package and ensembles code public when published
10. Sorrry - dumb questions, but probably should ask to properly understand equations... line 184 pdf - what is theta-hat? and what do you use it for? should you describe this, or is it obvious to more technically minded folk than me? I don't fully understand what b is. 
11. Also, why do you use 'arg min' and 'a loss function'? Am i right in thinking that the first equation is to find out which superensemble models are best predictors (different theta-hats) and the different 'loss functions' are those used to determine bias and accuracy? Or am I completely wrong? Would it help to explain what arg min and the loss functions are used for (as it took me a long time to figure out!)? you could just say 'for performance metrics, see below'
12. line 202 - is this cross validation of GBM the same as the cross-validation described in the section below about performance metrics?
13. FIgure 1 - I like it as is. 
14. Figure 4 - tweak labels when all else is completed.
15. Line 255 from pdf - CMSY predictions were approximately linear  (Fig S4) - Do you mean Fig S4a? Maybe I don't know how to interpret these graphs - I don't see how you can extract that info from the CMSY graph. I don't really know wht you mean by 'high and low prediction clusters'.
16. line 258-59 - if mPRM and SSCOM are high, then true status tends to be high. Is there another example that shows how covariates are exploited. The example shown is convincing in Fig S5, but not in Fig S6l, because the predicted B/Bmsy is only just above (light pink shading). though ths may not matter. you can ignore this comment if you want.
17. check for consistent US spelling - e.g. behaviour - line 279-280. there might be more that I didn't notice. also - line 290 and 291 - overfitting vs over-fitting. I'm not sure which is correct. 
18. line 314-15 need a reference showing that SSCOM contributed little to status from GBM. e.g. which figure shows this?
19. line 334 - should this be "effort dynamics at low biomasses in COM-SIR", not SSCOM? 
20. I said that i'd help with the last paragraph about other uses for ensemble methos. If there is space within the word limit, could we add just one other potential use? Are the references listed in 350-51 using ensemble models or do they simply provide examples of where ensembles could be used. Just check my addition - and delete it if you aren't happy with it. I don't know how to insert references - i'll leave that up to you. 
21. fix acknowledgements section
22. I didn't check references. 
23. other than that! Awesome! Hope my comments were helpful.

@@

We first developed and tested ensemble methods with a fully factorial simulated dataset [@rosenberg2014] with precisely known status. These published simulations included the following scenarios: three fish life histories: small pelagic, demersal, and large pelagic; three levels of how depleted biomass was at the start of the dataset compared to carrying capacity: 100%, 70%, and 40% of carrying capacity; and four harvest dynamics: a constant harvest rate, a harvest rate coupled with biomass to mimic an open-access single-species fishery, a scenario where harvest rate increases continuously, and a 'roller-coaster' scenario where the harvest rate increases and then decreases. Process noise (i.e. variability in population dynamics) was introduced to the models through two levels of multiplicative recruitment variability: $N(0, 0.2)$ and $N(0, 0.6)$, and was either uncorrelated or had a first-order autoregressive correlation of $0.6$. The simulation also included a scenario with multiplicative observation error around catch at $N(0, 0.2)$ and one without observation error. 
<!--TODO: make last sentence clearer-->

@@

We also tested our ensemble methods on the RAM Legacy Stock Assessment Database [@ricard2012]. Our analysis of the stock-assessment database was based on version XX downloaded on XX. After removing stocks for which the models described below did not converge, this database included `r ram_stocks_n` stocks.

@@

\noindent
where $\hat{B}_{t+1}$ represents predicted biomass at time $t$ plus one year, $B_t$ represents biomass at time $t$, $r$ represents the intrinsic population growth rate, $B_0$ represents unfished biomass or carrying capacity $K$, and $C$ represents catch. @rosenberg2014 provide a full summary of these four methods and code to fit all the models is available in an accompanying package `datalimited` for the statistical software \textsf{R}. In summary:

@@

\noindent 
where $L$ is a generalized loss function, $f$ is some generalized regression function, $\hat{b}_\mathrm{CMSY}$ is the mean or slope prediction from the CMSY model etc.\ (Table \ref{tab:predictors}), $S(0.2)$ and $S(0.05)$ represent the spectral density of the scaled catch series at frequencies of 0.20 and 0.05, $b$ is the mean or slope value the model is fitted to, and $\theta$ is the set of parameters for the regression function. For all superensemble models of mean $B/B_\mathrm{MSY}$ --- a ratio bounded at zero --- we fit the models in log space and exponentiated the predictions. For the estimates of $B/B_\mathrm{MSY}$ slope, which are not bounded at zero, we fit superensemble models on the natural untransformed scale.

We fit the linear model superensemble with second-order interactions. We fit the two machine learning superensemble models based on regression trees. Regression trees sequentially determine what value of a predictor best splits the response data into two 'branches' based on a specified loss function [@breiman1984]. In random forests, a series of regression trees are built on a random subset of the data and random subset of the covariates of the model [@breiman2001]. In generalized boosted models (GBMs), each subsequent model is fit to the residuals from the previous model; data points that are fit poorly in a given model are given more weight in the next model [@elith2008]. Random forests and GBMs can provide strong predictive performance and fit highly non-linear relationships [@elith2008; @hastie2009]. We fit random forest models with the `randomForest` package [@liaw2002] for \textsf{R} with the default argument values. We fit boosted regression tree models with the `gbm` package [@ridgeway2015] for \textsf{R}. Based on cross-validation with the **caret** \textsf{R} package, we fit GBMs with 2000 trees, an interaction depth of $6$, a learning rate (shrinkage parameter) of $0.01$, and all other arguments at their default values.

@@

Predictive performance can be evaluated with metrics that represent a variety of modelling goals. For continuous response variables such as the mean and slope of population status, performance metrics often measure some form of bias, precision, accuracy (a combination of bias and precision), or the ability to correctly rank or correlate across populations [e.g. @walther2005]. Here, we measure proportional error, defined as $(\hat{b} - b)/b$, where $\hat{b}$ and $b$ represent estimated and 'true' (or stock-assessed) mean or slope of $B/B_\mathrm{MSY}$. We calculated median proportional error to measure bias, median absolute proportional error to measure accuracy, and Spearman's rank-order correlation between predicted and 'true' values to measure the ability to correctly rank populations. We measured the ability for each model to correctly place stocks into categories of above or below $B/B_\mathrm{MSY} = 1$ at any decision threshold by plotting receiver operating characteristic curves. Code to reproduce our analysis is available at <https://github.com/datalimited/ensembles>.

@@

Ensemble methods provide a useful approach to situations where conservation and environmental resource management decisions must be made on the basis of multiple, potentially contrasting estimates of status --- a situation common to many settings beyond fisheries. Compared to individual models of fish population status, ensemble methods were consistently the best or among the best across three performance dimensions (accuracy, bias, and rank-order correlation), two response variables (status and trajectory), two datasets (simulated and global fisheries), and multiple ensemble methods (from a simple average to machine learning superensembles). Our results suggest choosing a superensemble model that allows for non-parametric relationships, such as machine learning methods. These methods provide added insight into individual model behaviour and generally performed the best; however, even a simple average of predictions across multiple models may provide a more useful measure of the desired parameter than a single model in many cases.

Certain conditions will make some ensemble models more effective than others. First, ensembles will be most effective when they are comprised of diverse individual models that choose different structural model forms, explore contrasting but plausible ranges of parameter values, or make errors in uncorrelated ways [@ali1996; @dietterich2000; @tebaldi2007]. Such individual models would be expected to perform well in different conditions and an ensemble model can exploit the best predictive performance of each. Second, ensemble models will be most effective when they are not overfit to the training dataset. Cross-validation testing [@caruana2004; @hastie2009] and methods that are robust to over-fitting such as random forests [@breiman2001], may help avoid overfitting ensemble models. We note that our simplest ensemble model, an average of individual model predictions, performed approximately as well as complex machine learning models when we trained our superensembles on the simulation dataset and tested them on a separate 'real' dataset (i.e. the RAM Legacy Stock Assessment database, Fig. \ref{performance}b). Third, ensemble models will be most effective when they are trained on data that are representative of the dataset of interest [@knutti2009; @weigel2010]. Cross-validation within a training dataset will provide an optimistically biased impression of predictive performance if the training dataset fundamentally differs from the dataset of interest [@hastie2009].

>>>>

<!--TODO: Jessicas Comments: 
15. Line 255 from pdf - CMSY predictions were approximately linear  (Fig S4) - Do you mean Fig S4a? Maybe I don't know how to interpret these graphs - I don't see how you can extract that info from the CMSY graph. I don't really know wht you mean by 'high and low prediction clusters'.
16. line 258-59 - if mPRM and SSCOM are high, then true status tends to be high. Is there another example that shows how covariates are exploited. The example shown is convincing in Fig S5, but not in Fig S6l, because the predicted B/Bmsy is only just above (light pink shading). though ths may not matter. you can ignore this comment if you want.
18. line 314-15 need a reference showing that SSCOM contributed little to status from GBM. e.g. which figure shows this?
19. line 334 - should this be "effort dynamics at low biomasses in COM-SIR", not SSCOM? 
20. I said that i'd help with the last paragraph about other uses for ensemble methos. If there is space within the word limit, could we add just one other potential use? Are the references listed in 350-51 using ensemble models or do they simply provide examples of where ensembles could be used. Just check my addition - and delete it if you aren't happy with it. I don't know how to insert references - i'll leave that up to you. 
-->

<!-- 
TODO
- justified or ragged?
- spell check American
-->

@@

We first developed and tested ensemble methods with a fully factorial simulated dataset [@rosenberg2014] with precisely known status. These published simulations included the following scenarios: three fish life histories: small pelagic, demersal, and large pelagic; three levels of how depleted biomass was at the start of the dataset compared to carrying capacity: 100%, 70%, and 40% of carrying capacity; and four harvest dynamics: a constant harvest rate, a harvest rate coupled with biomass to mimic an open-access single-species fishery, a scenario where harvest rate increases continuously, and a 'roller-coaster' scenario where the harvest rate increases and then decreases. Process noise (i.e. variability in population dynamics) was introduced to the models through two levels of multiplicative recruitment variability: $N(0, 0.2)$ and $N(0, 0.6)$, and was either uncorrelated or had a first-order autoregressive correlation of $0.6$. The simulation also included a scenario with multiplicative observation error around catch at $N(0, 0.2)$ and one without observation error. 
<!--TODO: make last sentence clearer
4. First paragraph in Methods describing simulations line 113 in pdf - you may want to describe the recrutiment variability a bit more. I know its not an important detail, but because its in the first paragraph of methods, it might turn people off reading further. Most people wouldn't understand what 'multiplcative' (including me) is or may not know what 'process error/noise' refers to. 
5. First paragraph in Methods last sentence - needed a bit of rewording because as was, i got quite confused. hopefully this new sentence makes sense and doesnt change the meaning. 
-->

@@

We also tested our ensemble methods on the RAM Legacy Stock Assessment Database [@ricard2012]. Our analysis of the stock-assessment database was based on version XX downloaded on XX (TODO). After removing stocks for which the models described below did not converge (TODO: add number), this database included `r ram_stocks_n` stocks.

@@

\noindent
where $\hat{B}_{t+1}$ represents predicted biomass at time $t$ plus one year, $B_t$ represents biomass at time $t$, $r$ represents the intrinsic population growth rate, $B_0$ represents unfished biomass or carrying capacity $K$, and $C$ represents catch. @rosenberg2014 provide a full summary of these four methods and code to fit all the models is available in an accompanying package `datalimited` (TODO: make public?) for the statistical software \textsf{R}. In summary:

@@

\noindent 
where $L$ is a generalized loss function, $f$ is some generalized regression function, $\hat{b}_\mathrm{CMSY}$ is the mean or slope prediction from the CMSY model etc.\ (Table \ref{tab:predictors}), $S(0.2)$ and $S(0.05)$ represent the spectral density of the scaled catch series at frequencies of 0.20 and 0.05, $b$ is the mean or slope value the model is fitted to, and $\theta$ is the set of parameters for the regression function. For all superensemble models of mean $B/B_\mathrm{MSY}$ --- a ratio bounded at zero --- we fit the models in log space and exponentiated the predictions. For the estimates of $B/B_\mathrm{MSY}$ slope, which are not bounded at zero, we fit superensemble models on the natural untransformed scale.
<!--add theta-hat description, b hat, make sure these are right, maybe simplify, describe arg min-->

We fit the linear model superensemble with second-order interactions. We fit the two machine learning superensemble models based on regression trees. Regression trees sequentially determine what value of a predictor best splits the response data into two 'branches' based on a specified loss function [@breiman1984]. In random forests, a series of regression trees are built on a random subset of the data and random subset of the covariates of the model [@breiman2001]. In generalized boosted models (GBMs), each subsequent model is fit to the residuals from the previous model; data points that are fit poorly in a given model are given more weight in the next model [@elith2008]. Random forests and GBMs can provide strong predictive performance and fit highly non-linear relationships [@elith2008; @hastie2009]. We fit random forest models with the `randomForest` package [@liaw2002] for \textsf{R} with the default argument values. We fit boosted regression tree models with the `gbm` package [@ridgeway2015] for \textsf{R}. Based on cross-validation with the **caret** \textsf{R} package, we fit GBMs with 2000 trees, an interaction depth of $6$, a learning rate (shrinkage parameter) of $0.01$, and all other arguments at their default values.                                                
<!--TODO give more details on tuning or gloss over more-->

@@

Predictive performance can be evaluated with metrics that represent a variety of modelling goals. For continuous response variables such as the mean and slope of population status, performance metrics often measure some form of bias, precision, accuracy (a combination of bias and precision), or the ability to correctly rank or correlate across populations [e.g. @walther2005]. Here, we measure proportional error, defined as $(\hat{b} - b)/b$, where $\hat{b}$ and $b$ represent estimated and 'true' (or stock-assessed) mean or slope of $B/B_\mathrm{MSY}$. We calculated median proportional error to measure bias, median absolute proportional error to measure accuracy, and Spearman's rank-order correlation between predicted and 'true' values to measure the ability to correctly rank populations. We measured the ability for each model to correctly place stocks into categories of above or below $B/B_\mathrm{MSY} = 1$ at any decision threshold by plotting receiver operating characteristic curves. Code to reproduce our analysis is available at <https://github.com/datalimited/ensembles> (TODO: make public).

@@

Ensemble methods provide a useful approach to situations where conservation and environmental resource management decisions must be made on the basis of multiple, potentially contrasting estimates of status --- a situation common to many settings beyond fisheries. Compared to individual models of fish population status, ensemble methods were consistently the best or among the best across three performance dimensions (accuracy, bias, and rank-order correlation), two response variables (status and trajectory), two datasets (simulated and global fisheries), and multiple ensemble methods (from a simple average to machine learning superensembles). Our results suggest choosing a superensemble model that allows for non-parametric relationships, such as machine learning methods. These methods provide added insight into individual model behavior and generally performed the best; however, even a simple average of predictions across multiple models may provide a more useful measure of the desired parameter than a single model in many cases.

Certain conditions will make some ensemble models more effective than others. First, ensembles will be most effective when they are comprised of diverse individual models that choose different structural model forms, explore contrasting but plausible ranges of parameter values, or make errors in uncorrelated ways [@ali1996; @dietterich2000; @tebaldi2007]. Such individual models would be expected to perform well in different conditions and an ensemble model can exploit the best predictive performance of each. Second, ensemble models will be most effective when they are not overfit to the training dataset. Cross-validation testing [@caruana2004; @hastie2009] and methods that are robust to overfitting such as random forests [@breiman2001], may help avoid overfitting ensemble models. We note that our simplest ensemble model, an average of individual model predictions, performed approximately as well as complex machine learning models when we trained our superensembles on the simulation dataset and tested them on a separate 'real' dataset (i.e. the RAM Legacy Stock Assessment database, Fig. \ref{performance}b). Third, ensemble models will be most effective when they are trained on data that are representative of the dataset of interest [@knutti2009; @weigel2010]. Cross-validation within a training dataset will provide an optimistically biased impression of predictive performance if the training dataset fundamentally differs from the dataset of interest [@hastie2009].
